### 2018-02-21 CO749-GraphColoring/assignments/A3.tex
Last claim of A3 remains

\section{}
\subsection{a}
\begin{lemma}
	If $G$ is the Ore-composition of two graphs $G_1$ and $G_2$ then $\alpha(G) \leq \alpha(G_1) + \alpha(G_2)$.
\end{lemma}
\begin{proof}
Let $G$ be the Ore-composition of graphs $G_1$ and $G_2$ obtained by deleting edge $xy \in E(G_1)$ and splitting vertex $z \in V(G_2)$ into $z_1$ and $z_2$ of positive degree and identifying $x$ with $z_1$ and $y$ with $z_2$. We denote the identified vertex for $x$ with $z_1$ as $z_x$, and for $y$ with $z_2$ as $z_y$ in $G$. Let $S \subseteq V(G)$ be an independent set in $G$ such that $|S| = \alpha (G)$. 
Then treating $z_x$ as $x$ and $z_y$ as $z$ we can partition $S$ into two independent sets $S_1\subseteq V(G_1)$ and $S_2\subseteq V(G_2)$ as $S_1 = V(G_1) \cap S$ and $S_2 = S\backslash S_1$. Thus
$$\alpha(G) = |S| = |S_1| + |S_2| \leq \alpha(G_1) + \alpha(G_2)$$
as desired.
\end{proof}

\subsection{b}
\begin{lemma}
	If $G$ is $k$-Ore then $\alpha(G) = \frac{v(G) - 1}{k-1}$.
\end{lemma}
\begin{proof}
	We proceed by induction on $v(G)$. The smallest $k$-Ore graph is $K_k$, for which
	$$\alpha(K_k) = 1 = \frac{k-1}{k-1} = \frac{v(G) -1}{k-1}.$$
	So the base case for induction is satisfied.
	\paragraph{}
	So we consider a $k$-Ore graph $G$ that is the Ore-composition of $G_1$ and $G_2$ and suppose for induction that
	$$\alpha(G_1) = \frac{v(G_1) - 1}{k-1} \quad\text{ and }\quad \alpha(G_2) = \frac{v(G_2) -1}{k-1}.$$
	We observe from the previous Lemma, and induction, that
	$$\alpha(G) \leq \alpha(G_1) + \alpha(G_2) = \frac{v(G_1) +v(G_2)- 2}{k-1} = \frac{v(G) + 1 - 2}{k-1} = \frac{v(G) - 1}{k-1}.$$
	Thus $\alpha(G) \leq \frac{v(G) - 1}{k-1}$. To show $\alpha(G) \geq \frac{v(G) - 1}{k-1}$ it suffices to construct an independent set in $G$ of size $\frac{v(G) - 1}{k-1}$.
	\paragraph{}
	For $i \in \{1, \dots, 2\}$ let $S_i$ be in an independent set in $G_i$ satisfying $$|S_i| = \alpha(G_i) = \frac{v(G_i) - 1}{k-1}.$$ We use the same notion as the above problems for $x,y, z, z_x, z_y$. Since $xy \in E(G_1)$, $|\{x,y\} \cap S_1| \leq 1$. Without loss of generality suppose $y \not\in S_1$. Treating $x$ as $z_x$ and $z$ as $z_y$ we observe that
	$$S := S_1 \cup S_2$$
	is an independent set of $G$ of size $|S_1| + |S_2|$. But then we see that
	$$|S| = |S_1| + |S_2| =  \frac{v(G_1) +v(G_2)- 2}{k-1} = \frac{v(G) - 1}{k-1}.$$
	Thus $\alpha(G) \geq \frac{v(G) - 1}{k-1}$, and therefore by induction the Lemma is proven.
\end{proof}
\end{document}

>>>>

\newpage
\section{}
\subsection{a}
\begin{lemma}
	If $G$ is the Ore-composition of two graphs $G_1$ and $G_2$ then $\alpha(G) \leq \alpha(G_1) + \alpha(G_2)$.
\end{lemma}
\begin{proof}
Let $G$ be the Ore-composition of graphs $G_1$ and $G_2$ obtained by deleting edge $xy \in E(G_1)$ and splitting vertex $z \in V(G_2)$ into $z_1$ and $z_2$ of positive degree and identifying $x$ with $z_1$ and $y$ with $z_2$. We denote the identified vertex for $x$ with $z_1$ as $z_x$, and for $y$ with $z_2$ as $z_y$ in $G$. Let $S \subseteq V(G)$ be an independent set in $G$ such that $|S| = \alpha (G)$. 
Then treating $z_x$ as $x$ and $z_y$ as $z$ we can partition $S$ into two independent sets $S_1\subseteq V(G_1)$ and $S_2\subseteq V(G_2)$ as $S_1 = V(G_1) \cap S$ and $S_2 = S\backslash S_1$. Thus
$$\alpha(G) = |S| = |S_1| + |S_2| \leq \alpha(G_1) + \alpha(G_2)$$
as desired.
\end{proof}

\subsection{b}
\begin{lemma}
	If $G$ is $k$-Ore then $\alpha(G) = \frac{v(G) - 1}{k-1}$.
\end{lemma}
\begin{proof}
	We proceed by induction on $v(G)$. The smallest $k$-Ore graph is $K_k$, for which
	$$\alpha(K_k) = 1 = \frac{k-1}{k-1} = \frac{v(G) -1}{k-1}.$$
	So the base case for induction is satisfied.
	\paragraph{}
	So we consider a $k$-Ore graph $G$ that is the Ore-composition of $G_1$ and $G_2$ and suppose for induction that
	$$\alpha(G_1) = \frac{v(G_1) - 1}{k-1} \quad\text{ and }\quad \alpha(G_2) = \frac{v(G_2) -1}{k-1}.$$
	We observe from the previous Lemma, and induction, that
	$$\alpha(G) \leq \alpha(G_1) + \alpha(G_2) = \frac{v(G_1) +v(G_2)- 2}{k-1} = \frac{v(G) + 1 - 2}{k-1} = \frac{v(G) - 1}{k-1}.$$
	Thus $\alpha(G) \leq \frac{v(G) - 1}{k-1}$. To show $\alpha(G) \geq \frac{v(G) - 1}{k-1}$ it suffices to construct an independent set in $G$ of size $\frac{v(G) - 1}{k-1}$.
	\paragraph{}
	For $i \in \{1, \dots, 2\}$ let $S_i$ be in an independent set in $G_i$ satisfying $$|S_i| = \alpha(G_i) = \frac{v(G_i) - 1}{k-1}.$$ We use the same notion as the above problems for $x,y, z, z_x, z_y$. Since $xy \in E(G_1)$, $|\{x,y\} \cap S_1| \leq 1$. Without loss of generality suppose $y \not\in S_1$. Treating $x$ as $z_x$ and $z$ as $z_y$ we observe that
	$$S := S_1 \cup S_2$$
	is an independent set of $G$ of size $|S_1| + |S_2|$. But then we see that
	$$|S| = |S_1| + |S_2| =  \frac{v(G_1) +v(G_2)- 2}{k-1} = \frac{v(G) - 1}{k-1}.$$
	Thus $\alpha(G) \geq \frac{v(G) - 1}{k-1}$, and therefore by induction the Lemma is proven.
\end{proof}

\subsection{c}
From our work in class it suffices to show $c\alpha(R') \leq c\alpha(R) + c\alpha(W) - c\alpha(K_{|X|}) + c$ for $c \geq 0$. Since we know the core is always non-empty, $\alpha(K_{|X|}) = 1$ and so the problem reduces to showing
$$\alpha(R') \leq \alpha(R) + \alpha(W).$$
Let $S$ be an independent set of $G[R']$ such that $|S| = \alpha(R')$. Since $R'= V(W) \cup R - \{v_1, \dots, v_{k-1}\}$, where $\{v_1, \dots, v_{k-1}\}$ are the set of color-identified vertices, as in the usual definition. So we have
$$|S| \leq |S\cap (V(W) - \{v_1, \dots, v_k\}| + |S\cap R|$$
On the one hand $S\cap (V(W) - \{v_1, \dots, v_k\})$ is an independent set in $W$ and $S \cap R$ is an indepenent set in $G[R]$, so then
$$ \alpha(R') = |S| \leq |S\cap (V(W) - \{v_1, \dots, v_k\}| + |S\cap R| \leq \alpha(W) + \alpha(R)$$
as desired. $\blacksquare$
\newpage
\subsection{a}
If $G$ is the Ore-composition of $G_1$ and $G_2$ then
$$T_k(G) \geq T_k(G_1-e) + T_k(G_2 - z) \geq T_k(G_1) + T_k(G_2) - 2$$
where $e = xy$ is the replaced edge of $G_1$ and $z$ is the split vertex of $G_2$.
Let $H_1 \subseteq G_1- e$ be the subgraph of $G_1 - e$ which achieves the optimum in 
$$T_k(G_1-e) := \max_{H \subseteq G_1 - e} T_k(H).$$
Similarly let $H_2 \subseteq G_2 - z$ be the subgraph of $G_2 - z$ which achieves the optimum in $T_k(G_2 - z)$. Then $H_1 \cup H_2$ is a disjoint union of cliques in $G$. So we have
$$T_k(G_1-e) + T_k(G_2 - z) = T_k(H_1) + T_k(H_2) = T_k(H_1 \cup H_2) \leq T_k(G).$$
So the first inequality of the lemma is proven.
Now let $H_1' \subseteq G_1$ be the subgraph of $G_1$ which achieves the optimum in $T_k(G_1)$, and let $H_2' \subseteq G_2$ be the subgraph of $G_2$ which achieves the optimum in $T_k(G_2)$. If $|e \cap V(H'_1)| \leq 1$ then $H'_1$ is a subgraph of $G_1 - e$ and so $T_k(G_1) \leq T_k(G_1-e)$. Otherwise, i.e. $ e\in E(H'_1)$ then $H'_1 - x$ is a subgraph of $G_1 - e$, and $T_k(H'_1-x) = T_k(H'_1) - 1$. Thus 
$$T_k(G_1 - e) \geq T_k(H'_1-x) \geq T_k(H'_1) - 1 =T_k(G_1) - 1.$$
Similarly $H'_2 - z$ is a subgraph of $G_2 - v$ and $T_k(H'_2-z) \geq T_k(H'_2) - 1$. So 
$$T_k(G_2 - z) \geq T_k(H'_2 - z) \geq T_k(H'_2) - 1 = T_k(G_2) - 1.$$
Summing our inequalities above we obtain
$$T_k(G_1 - e) + T_k(G_2 - z) \geq T_k(G_1) + T_k(G_2) - 2$$
as desired.
\subsection{b}
If $G$ is $k$-Ore and $G \neq K_k$ then $T_k(G) \geq 2+ \frac{v(G) - 1}{k-1}$.
Suppose not and let $G$ be a minimum counterexample. Since $G$ is $k$-Ore and $G\neq K_k$ there exist $k$-Ore graphs $G_1$ and $G_2$ such that $G$ is the Ore-composition of $G_1$ and $G_2$. We proceed by case distinction on if $G_1, G_2$ are $K_k$ or not. If neither $G_1$ nor $G_2$ are $K_k$ then by minimality of $G$,
$$T_k(G_i) \geq 2 + \frac{v(G_i) - 1}{k-1}$$
for $i =1,2$. But then combining with the inequality
$$T_k(G) \geq T_k(G_1) + T_k(G_2) - 2$$
from the previous Lemma, we see that
$$T_k(G) \geq 2 + \frac{v(G_1) - 1}{k-1} + 2 + \frac{v(G_2) - 1}{k-1} -2 \geq 2 + \frac{v(G)-1}{k-1}$$
as desired, using that $v(G_1) + v(G_2) -1 = v(G)$.
If $G_1 = K_k$ and $G_2 \neq K_k$ then using the first inequality from the previous Lemma we have
$$T_k(G) \geq T_k(G_1-e) + T_k(G_2-z) \geq 2 + T_k(G_2-z)$$
Since $T_k(G_2-z) \geq T_k(G_2) - 1$, 
$$T_k(G) \geq 1 + T_k(G_2) \geq \frac{k-1}{k-1} + \frac{v(G_2)-1}{k-1} + 2= \frac{v(G_1) + v(G_2) - 2}{k-1} +2 = \frac{v(G) - 1}{k-1} +2$$
The proof for $G_2 = K_k$ and $G_1 \neq K_k$ follows similarly.
If $G_1 = K_k$ and $G_2 = K_k$ then 
$$T_k(G) \geq T_k(G_1-e) + T_k(G_2 - z) = 2  + 2 = \frac{2(k-1)}{k-1} + 2 = \frac{v(G_1) + v(G_2) -2}{k-1} + 2 = \frac{v(G) -1}{k-1} + 2$$
as desired.
\subsection{c}
It suffices to show
$$T_k(R) + T_k(W) - T_k(K_{|X|}) \leq T_k(R') + |X|.$$
Let $H_R \subseteq G[R]$ be a disjoint union of cliques attaining the maximum in $T_k(R)$. Let $H_W \subseteq W$ be a disjoint union of cliques attaining the maximum in $T_k(W)$. If a vertex in $X$ is deleted from a clique in $H_W$ then the value of $T_k$ for the resulting graph will decrease by at most one. Hence $T_k(H_W - X) \geq T_k(H_W) - |X|$. But then $H_R \cup (H_W - X)$ is a disjoint union of cliques in $G[R']$. So we have
\begin{align*}
T_k(R) + T_k(W) - T_k(K_{|X|}) &\leq T_k(H_R) + T_k(H_W)\\ &\leq T_k(H_R) + T_k(H_W - X) - |X|\\ &\leq T_k(H_R \cup (H_W-X)) - |X|\\ &\leq T_k(R') - |X|\end{align*}
as desired. $\blacksquare$
\newpage
Let $k \geq 4$. Let $p(G) = (k-1)v(G) - 2e(G)$. Then $p(G) \leq 0$ if $G$ is $k$-critical since the minimum degree of $G$ is $k-1$. Let $G$ be a counterexample to the potential version of Brooks' Theorem. That is, $G$ is $k$-critical, $G\neq K_k$, and $p(G) = 0$. Since $p(G) = 0$, we have that $G$ is $(k-1)$-regular.
\begin{claim}
If $R \subset V(G)$ then $p(R) \geq k-1$.
\end{claim}
Suppose not and let $R$ be a counterexample with maximum number of vertices. Then $R$ has an extension $R'$ with extender $W$ and core $X$. By the potential-extension lemma,
$$p(R') \leq p(R) + p(W) - p(K_{|X|})$$
Now $$p(R) =  (k-1)|R| - 2e(R) \geq (k-|R|)|R|$$
which is at least $k-1$ when $1 \leq |R| \leq k-1$. Hence $|R| \geq k$, and since $G$ is not equal to $K_k$ and $G$ is $(k-1)$-regular, $G[R]$ is not a clique of size at least $k$. Therefore $W$ is a strict subgraph of $G$, and so by minimality of $G$,
$$p(W) \leq -1.$$
By the same reasoning as above for $R$, $$p(K_{|X|}) \geq k-1.$$
Thus we have
$$p(R') \leq p(R) + p(W) - p(K_{|X|}) \leq (k-2)  - 1 -(k-1) \leq - 2$$
By maximality of $R$, $p(R') \geq k-1$ contradicting the above.
\begin{claim}
If $z$ is a vertex with two non-adjacent neighbours $x,y$ then there exists a $(k-1)$-clique $H$ in $G\backslash \{x,y,z\}$ such that $N(x) \cap V(H)$, $N(y) \cap V(H)$ partition $V(H)$.
\end{claim}

\end{proof}
\end{document}
